---
title: "Applying ACM0009 with cdmAcm0009"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Applying ACM0009 with cdmAcm0009}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
```

# Overview

The `cdmAcm0009` package implements the essential monitoring and greenhouse gas accounting logic for the CDM large-scale methodology ACM0009. The methodology covers projects that switch existing power or heat plants from coal and/or petroleum fuels to natural gas.

# Workflow

1. Validate that the project meets the applicability conditions, including a qualifying baseline fuel mix and natural gas as the project fuel.
2. Calculate baseline and project emissions using fuel consumption, net calorific values, and emission factors.
3. Quantify leakage, including methane slip from gas handling and other project-specific sources.
4. Aggregate period-level results to report total emission reductions.

```{r}
library(cdmAcm0009)

data <- simulate_acm0009_dataset(periods = 3, seed = 12)
results <- estimate_emission_reductions_acm0009(data)
results$total_emission_reductions
```
