---
title: "cdmAmsIIIe"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# cdmAmsIIIe

`cdmAmsIIIe` implements the Clean Development Mechanism (CDM) small-scale methodology **AMS-III.E Avoidance of methane production from decay of biomass (combustion / gasification / mechanical-thermal treatment)**. The package exposes applicability diagnostics, equation-level helpers, simulation utilities, and orchestration wrappers that align with tidyverse data practices.

## Installation

```r
# install.packages("devtools")
devtools::install_local("path/to/cdmAmsIIIe")
```

## Example

```{r example, message = FALSE}
library(cdmAmsIIIe)
library(dplyr)

simulated <- simulate_ams_iiie_dataset(n_plants = 2)

applicability <- simulated$applicability |> 
  check_applicability_feedstock_characteristics_iiie(group_cols = "plant_id") |> 
  dplyr::left_join(
    check_applicability_biomass_control_iiie(simulated$applicability, group_cols = "plant_id"),
    by = "plant_id"
  ) |> 
  dplyr::left_join(
    check_applicability_monitoring_practices_iiie(simulated$applicability, group_cols = "plant_id"),
    by = "plant_id"
  )

workflow <- estimate_emission_reductions_ams_iiie(
  baseline_data = simulated$baseline,
  project_data = simulated$project,
  leakage_data = simulated$leakage,
  group_cols = c("plant_id")
)

applicability
workflow
```
